Prediction of Periprosthetic Joint Infection after Total Hip Arthroplasty: a simplified and externally validated model based on observational registry data from Sweden and ...

### 2020-09-30 reports/article.Rmd
include validation template

@@

**PARTICIPANTS AND METHODS: ** We analysed `r nmbr(N_shar)` patients recorded in the Swedish hip arthroplasty register `r period` and obtained data on co-morbidities and socioeconomic background variables from the National Patient Register and Statistics Sweden. Logistic least absolute shrinkage and selection operator (LASSO) regression was applied to develop prediction models for PJI within 90 days or two years after surgery. Models were internally validated by bootstrapping and externally validated on a cohort of **XXX** patients from **XXX**. We assessed model performance by analysing the area under the receiver operation characteristics curve (AUC) and by calibration belt plots.

**RESULTS: ** The crude incidence rate of PJI after THA was `r crude_rate_90d * 100` % within 90, days and `r crude_rate_2y * 100` % within two years in the Swedish derivation cohort. A prediction model for PJI within 90 days combined `r coefs_text["90 days"]`, resulting in an `r get_auc("90 days", "Reduced model", "2")` internally  and **(AUC=XXX, 95% CI: XXX to XXX)** externally. A model for PJI within two years had similar performance internally `r get_auc("2 years", "Reduced model")` and externally **XXX (95% CI: XXX to XXX)**. These models were superior to traditional models based on the American Society for Anesthesiologists (ASA) classification `r get_auc("90 days", "ASA class")`, Charlson `r get_auc("90 days", "Charlson")`, Elixhauser `r get_auc("90 days", "Elixhauser")`, and the Rx Risk V `r get_auc("90 days", "Rx Risk V")` co-morbidity indices.

@@

Periprosthetic joint infection (PJI) is the most devastating of all early complications after total hip arthroplasty (THA), mainly due to its severity in terms of increased mortality, the required number of re-operations, long-term antibiotic treatment, and often persistently impaired quality of life.[@Zimmerli2006; @Kapadia2016; @Shohat2019] Many risk factors for the development of PJI are identified, with anemia, diabetes, and obesity being important, to some extent modifiable risk factors, whereas advanced age and male gender are examples of non-modifiable risk factors.[@Ridgeway2005; @Lai2007; @Malinzak2009; @Ong2009; @Pedersen2010; @Bozic2012; @Jamsen2012; @Rasouli2014a; @Maoz2015]
Individualized PJI risk stratification would be a highly desirable tool in the pre-operative assessment of patients scheduled for THA, but established measures of co-morbidity such as the American Society for Anesthesiologists (ASA) classification or considerably more complex co-morbidity indices cannot be used in this context. Although high ASA class is a risk factor associated with development of PJI, the classification is imprecise, and it was never developed for the purpose of predicting this specific complication.[@Haynes1995; @Berbari2012] 
The Charlson[@Charlson1987] and Elixhauser[@Elixhauser1998] co-morbidity indices are based on `r nrow(coder::charlson)` or `r nrow(coder::elixhauser)` conditions coded by the International Classification of Diseases (ICD), respectively, but they are cumbersome to use in a clinical setting, and they were also not designed for prediction of specific adverse events after THA. The same is true for the Rx Risk V-classification with `r nrow(coder::rxriskv)` conditions codified by the Anatomical Therapeutic Chemical (ATC) Classification.[@Sloan2003; @Pratt2018]
Several attempts have been made at developing individualized risk prediction models for clinical usage, with the Universal American College of Surgeons National Surgical Quality Improvement Program (ACS NSQIP) Surgical Risk Calculator being one of the first.[@Edelstein2015; @Wingert2016a] Several other risk calculators for the risk of PJI after THA have since been derived on US or Australian cohorts, but none has gained general acceptance.[@Wingert2016a; @Bozic2013a; @Inacio2015a] This is mainly due to often limited, sometimes single-center based samples, low  discriminative ability or low accuracy of the prediction models, or the absence of external model validation.
Thus, there is an urgent need for improved prediction models of PJI. We therefore developed a prediction model of PJI within 90 days or two years after THA that is based upon easily accessible data that are available in the setting of clinical decision-making.

@@

The derivation cohort was obtained from patients recorded in the SHAR `r period`. The starting point was chosen because registration of ASA class and body mass index (BMI) in the SHAR was initiated in 2008, and we had no access to data from the national patient register (NPR) after 2015. Data linkage was achieved by using personal identity numbers that are assigned to all Swedish residents at birth or immigration.[@Ludvigsson2009; @Cnudde2016] The diagnosis underlying the index THA procedure (categorized into `r glue::glue_collapse(tolower(levels(df$P_DiaGrp)), ", ", last = " and ") %>% {gsub("\\(avn\\)", "[AVN]", .)}`), type of fixation (cemented or uncemented stem and cup respectively), age at index surgery, sex, BMI, ASA class, type of hospital, and year of surgery, were obtained from the SHAR. Data on co-morbidities for one year prior to index surgery were obtained from the NPR where diagnoses based on in- and outpatient episodes in all private and public hospitals are coded according to ICD-10 together with admission and discharge dates. Data on marital status and the educational level  of patients was obtained from the longitudinal integration database for health insurance and labor market studies (LISA) from Statistics Sweden. The SHAR has completeness of 96-98 % and 100 % coverage,[@Karrholm2018] the LISA has coverage of 98 % with 85 % accuracy, and completeness for the NPR is above 99 % and the positive predictive value is 85-95 %.[@Ludvigsson2011] Death dates were linked from the national population register. Only the last operated hip was considered for patients with bilateral THA,[@Bulow2020] and only patients who survivied at least two years aftert surgery. Patients with missing information on, or an ASA class of IV and above, were excluded, as were patients with missing information on BMI or a BMI above 50. Patients with missing information on educational levels or the type of hospital performing index surgery were also excluded.

@@

## Definitions of co-morbidity and outcomes

The Charlson, Elixhauser and Rx Risk V co-morbidity indices are based on the presence of ICD-10 and ATC codes grouped into main categories (`r nrow(coder::charlson)`, `r nrow(coder::elixhauser)` and `r nrow(coder::rxriskv)` respectively).[@Quan2005; @Pratt2018] Diagnostic categories from all indices were combined to define `r nrow(tab_categorization)` categories of co-morbidity that were chosen in order to be easily identifiable in a clinical setting (Tab. \@ref(tab:tabcategorization)). Some of the combined co-morbidities were observed for less than 10 patients with or without PJI. These diagnoses were not considered as potential predictors to reduce the risk of overfitting to spurious events. PJI was defined by the occurrence of relevant ICD-10 or procedural NOMESCO-code recorded in NPR within 90 days or 2 years (Tab. \@ref(tab:infcodes)), or if the patient was reported to SHAR, within the same time frame, as reoperated due to infection.

@@

Logistic LASSO regression was performed for each sample using the chosen $\lambda_i$ as penalty term. Regression coefficients were ranked based on their absolute values as a measure of variable importance. We then used Bayesian methods to estimate a posterior distribution and to identify two change points in a linear segmented piecewise regression model. The second change point was then used as a break-point. Influential predictors with absolute coefficients value above this point, scored a point. The whole process, each with 100 Bootstrap replicates, was repeated 100 times (yielding 10,000 bootstrap samples in total). Potential predictors that were selected at least once out of 100 were then used in main effects multivariable logistic regression models without penalty and without pre-normalization of numeric variables. Even more parsimonious models with potential predictors chosen at least 50 out of 100 times were evaluated for comparison. Univariable prediction models based on either ASA class, the Charlson, Elixhauser, or the Rx Risk V co-morbidity indices were fitted for comparison. Odds ratios for the final models were estimated with 95 % confidence intervals.

@@

We included `r nmbr(N_shar)` patients (`r mean(df$P_Sex == "Male") * 100` % males) with a mean age of `r age` years in the derivation cohort from the SHAR (Fig. \@ref(fig:flowchart), left panel). 
`r crude_rate_90d * 100` % (n = `r sum(df$outcome_infection_90d)`) of the patients developed a PJI within 90 days, and `r crude_rate_2y * 100` % (n = `r sum(df$outcome_infection_2y)`) within 2 years (Tab. \@ref(tab:tab1)). The proportion of patients with ASA class III was
`r comorb(df$P_ASA == "III")`; `r comorb(df$CCI_index_quan_original > 0)` had at least one comorbidity according to the Charlson co-morbidity index, `r comorb(df$ECI_index_sum_all > 0)` according to Elixhauser, and `r comorb(df$Rx_index_pratt > 0)` according to Rx Risk V. In addition, **XXX** patients from **XXX** were included as an external validation cohort (Fig. \@ref(fig:flowchart), right panel).

@@

Patients with `r excl_factors` were uncommon, wherefore those co-morbidities were excluded as potential predictors prior to further modeling. ROC-curves (Fig. \@ref(fig:rocs)) and AUC-values (Fig. \@ref(fig:aucci)) for the main and reduced models were no different. For simplicity, we therefore focus on the reduced models. Estimated probability density curves for patients with and without PJI were partially overlapping, although patients with PJI had on average higher predicted probabilities for such event (Fig. \@ref(fig:sep)).

The reduced model for the prediction of PJI within 90 days in the Swedish cohort included `r coefs_text["90 days"]` (Tab. \@ref(tab:coefs)). Predictors associated with a statistically significant risk increase were **secondary osteoarthritis as the underlying reason for THA surgery, male gender, obesity, and the presence of psoriasis, osteoporosis, Paget's or liver disease**. This reduced model for the predicition of PJI within 90 days had an `r get_auc("90 days", "Reduced model", "2")`, whereas ASA class `r get_auc("90 days", "ASA class")`, the Charlson co-morbidity index `r get_auc("90 days", "Charlson")`, the Elixhauser co-morbidity index `r get_auc("90 days", "Elixhauser")`, and Rx Risk V `r get_auc("90 days", "Rx Risk V")` had less discriminative ability (Fig. \@ref(fig:rocs) and \@ref(fig:aucci)). The reduced model had good accuracy as seen by the calibration plot (Fig. \@ref(fig:calibration), left panel).

The reduced model for PJI within two years included the same predictors, with the addition of `r coefs_text["2 years"]` (Tab. \@ref(tab:coefs)). Predictors conferring a statistically significant risk increase were those **mentioned as predictive for PJI within 90 days, excluding osteoporosis or Paget's disease, and additionally included an underlying diagnosis of AVN or inflammatory joint disease, and the presence of  CNS disease**. The reduced model for the prediction of PJI within two years had an `r get_auc("2 years", "Reduced model", "2")`, whereas ASA class `r get_auc("2 years", "ASA class")`, the Charlson co-morbidity index `r get_auc("2 years", "Charlson")`, the Elixhauser co-morbidity index `r get_auc("90 days", "Elixhauser")`, and Rx Risk V `r get_auc("90 days", "Rx Risk V")` again had less discriminative ability (Fig. \@ref(fig:rocs) and \@ref(fig:aucci)). Calibration was good for the reduced model (Fig. \@ref(fig:calibration), right panel).

@@

We found that a multivariable main effects logistic regression model based on some easily identifiable clinical conditions, gender, and BMI was considerably better at predicting PJI within 90 days after THA than models based on either ASA class or the Charlson, Elixhauser or Rx Risk V co-morbidity indices. A similarly developed model to predict PJI within two years included similar predictors as the model developed to predict PJI within 90 days, but added `r coefs_text["2 years"]` to the list of predictor variables. This model also performed better than the established co-morbidity measures in terms of AUC and calibration.

@@

Some previously mentioned risk factors for developing PJI were not selected during our model development. High ASA class is frequently associated with an increased risk of PJI,[@Berbari2012; @Pulido2008; @Namba2012] as are co-morbidity in a broader sense, and, more specifically, the presence of cardiovascular co-morbidity.[@Ong2009; @Higuera2011] In contrast, we found that ASA class alone was an insufficient predictor of PJI, and none of the cardiovascular disease categories remained as a predictor variable in our final main model.

Cancer seems to confer an increased risk of PJI in several observational studies,[@Bozic2012; @Everhart2013; @Poultsides2013] and it is one of the defined co-morbidities in the ACS NSQIP Surgical Risk Calculator,[@Wingert2016a]
20 but this comorbidity did not reach statistical significance in our models.

@@

Although mentioned as risk factors for surgical site infections after joint arthroplasty,[@Rasouli2014a; @Inacio2015a] prediction models based on the co-morbidity indices developed by Charlson or Elixhauser and the RxRisk V index alone, resulted in poor predictive power. This agrees with previous findings describing that these measures of comorbidity rather poorly predict mortality and patient reported outcomes after THA.[@Bulow2017; @Bulow2019]

@@

```{r aucci, fig.cap = "Area Under the Receiver Operation Characteristics Curve (AUC) as a measure of predictive discriminative ability with 95 % bootstrap confidence intervals. The reduced model performs no different than the main model, but both of these perform better than traditional co-morbidity measures alone."}

@@

tab_categorization %>%
  ungroup() %>% 
  mutate(
    `Comorbidities by groups` = 
      gsub("Heart infarct", "Myocardial infarction", `Comorbidities by groups`)
  ) %>% 
  knitr::kable(
    caption = "Categorization of co-morbidities from the Charlson, Elixhauser and Rx RIsk V co-morbidities. (CNS = central nervous system.)")

>>>>

# Note!

MOst text in the article is automatically generated! Bold text is not!
It is thus possible that this text does not correspond to the rest!
It must be updated manually!

@@

**PARTICIPANTS AND METHODS: ** We analysed `r nmbr(N_shar)` patients recorded in the Swedish hip arthroplasty register `r period` and obtained data on comorbidities and socioeconomic background variables from the National Patient Register and Statistics Sweden. Logistic least absolute shrinkage and selection operator (LASSO) regression was applied to develop prediction models for PJI within 90 days or two years after surgery. Models were internally validated by bootstrapping and externally validated on a cohort of **XXX** patients from **XXX**. We assessed model performance by analysing the area under the receiver operation characteristics curve (AUC) and by calibration belt plots.

**RESULTS: ** The crude incidence rate of PJI after THA was `r crude_rate_90d * 100` % within 90, days and `r crude_rate_2y * 100` % within two years in the Swedish derivation cohort. A prediction model for PJI within 90 days combined `r coefs_text["90 days"]`, resulting in an `r get_auc("90 days", "Reduced model", "2")` internally  and **(AUC=XXX, 95% CI: XXX to XXX)** externally. A model for PJI within two years had similar performance internally `r get_auc("2 years", "Reduced model")` and externally **XXX (95% CI: XXX to XXX)**. These models were superior to traditional models based on the American Society for Anesthesiologists (ASA) classification `r get_auc("90 days", "ASA class")`, Charlson `r get_auc("90 days", "Charlson")`, Elixhauser `r get_auc("90 days", "Elixhauser")`, and the Rx Risk V `r get_auc("90 days", "Rx Risk V")` comorbidity indices.

@@

Periprosthetic joint infection (PJI) is the most devastating of all early complications after total hip arthroplasty (THA), mainly due to its severity in terms of increased mortality, the required number of re-operations, long-term antibiotic treatment, and often persistently impaired quality of life.[@Zimmerli2006; @Kapadia2016; @Shohat2019] Many risk factors for the development of PJI are identified, with anemia, diabetes, and obesity being important, to some extent modifiable risk factors, whereas advanced age and male gender are examples of non-modifiable risk factors.[@Ridgeway2005; @Lai2007; @Malinzak2009; @Ong2009; @Pedersen2010; @Bozic2012; @Jamsen2012; @Rasouli2014a; @Maoz2015]
Individualized PJI risk stratification would be a highly desirable tool in the pre-operative assessment of patients scheduled for THA, but established measures of comorbidity such as the American Society for Anesthesiologists (ASA) classification or considerably more complex comorbidity indices cannot be used in this context. Although high ASA class is a risk factor associated with development of PJI, the classification is imprecise, and it was never developed for the purpose of predicting this specific complication.[@Haynes1995; @Berbari2012] 
The Charlson[@Charlson1987] and Elixhauser[@Elixhauser1998] comorbidity indices are based on `r nrow(coder::charlson)` or `r nrow(coder::elixhauser)` conditions coded by the International Classification of Diseases (ICD), respectively, but they are cumbersome to use in a clinical setting, and they were also not designed for prediction of specific adverse events after THA. The same is true for the Rx Risk V-classification with `r nrow(coder::rxriskv)` conditions codified by the Anatomical Therapeutic Chemical (ATC) Classification.[@Sloan2003; @Pratt2018]
Several attempts have been made at developing individualized risk prediction models for clinical usage, with the Universal American College of Surgeons National Surgical Quality Improvement Program (ACS NSQIP) Surgical Risk Calculator being one of the first.[@Edelstein2015; @Wingert2016a] Several other risk calculators for the risk of PJI after THA have since been derived on US or Australian cohorts, but none has gained general acceptance.[@Wingert2016a; @Bozic2013a; @Inacio2015a] This is mainly due to often limited, sometimes single-center based samples, low  discriminative ability or low accuracy of the prediction models, or the absence of external model validation.
Thus, there is an urgent need for improved prediction models of PJI. We therefore developed a prediction model of PJI within 90 days or two years after THA that is based upon easily accessible data that are available in the setting of clinical decision-making.

@@

The derivation cohort was obtained from patients recorded in the SHAR `r period`. The starting point was chosen because registration of ASA class and body mass index (BMI) in the SHAR was initiated in 2008, and we had no access to data from the national patient register (NPR) after 2015. Data linkage was achieved by using personal identity numbers that are assigned to all Swedish residents at birth or immigration.[@Ludvigsson2009; @Cnudde2016] The diagnosis underlying the index THA procedure (categorized into `r glue::glue_collapse(tolower(levels(df$P_DiaGrp)), ", ", last = " and ") %>% {gsub("\\(avn\\)", "[AVN]", .)}`), type of fixation (cemented or uncemented stem and cup respectively), age at index surgery, sex, BMI, ASA class, type of hospital, and year of surgery, were obtained from the SHAR. Data on comorbidities for one year prior to index surgery were obtained from the NPR where diagnoses based on in- and outpatient episodes in all private and public hospitals are coded according to ICD-10 together with admission and discharge dates. Data on marital status and the educational level  of patients was obtained from the longitudinal integration database for health insurance and labor market studies (LISA) from Statistics Sweden. The SHAR has completeness of 96-98 % and 100 % coverage,[@Karrholm2018] the LISA has coverage of 98 % with 85 % accuracy, and completeness for the NPR is above 99 % and the positive predictive value is 85-95 %.[@Ludvigsson2011] Death dates were linked from the national population register. Only the last operated hip was considered for patients with bilateral THA.[@Bulow2020]. Patients with missing information on, or an ASA class of IV and above, were excluded, as were patients with missing information on BMI or a BMI above 50. Patients with missing information on educational levels or the type of hospital performing index surgery were also excluded.

@@

## Definitions of comorbidity and outcomes

The Charlson, Elixhauser and Rx Risk V comorbidity indices are based on the presence of ICD-10 and ATC codes grouped into main categories (`r nrow(coder::charlson)`, `r nrow(coder::elixhauser)` and `r nrow(coder::rxriskv)` respectively).[@Quan2005; @Pratt2018] Diagnostic categories from all indices were combined to define `r nrow(tab_categorization)` categories of comorbidity that were chosen in order to be easily identifiable in a clinical setting (Tab. \@ref(tab:tabcategorization)). Some of the combined comorbidities were observed for less than 10 patients with or without PJI. These diagnoses were not considered as potential predictors to reduce the risk of overfitting to spurious events. PJI was defined by the occurrence of relevant ICD-10 or procedural NOMESCO-code recorded in NPR within 90 days or 2 years (Tab. \@ref(tab:infcodes)), or if the patient was reported to SHAR, within the same time frame, as reoperated due to infection.

@@

Logistic LASSO regression was performed for each sample using the chosen $\lambda_i$ as penalty term. Regression coefficients were ranked based on their absolute values as a measure of variable importance. We then used Bayesian methods to estimate a posterior distribution and to identify two change points in a linear segmented piecewise regression model. The second change point was then used as a break-point. Influential predictors with absolute coefficients value above this point, scored a point. The whole process, each with 100 Bootstrap replicates, was repeated 100 times (yielding 10,000 bootstrap samples in total). Potential predictors that were selected at least once out of 100 were then used in main effects multivariable logistic regression models without penalty and without pre-normalization of numeric variables. Even more parsimonious models with potential predictors chosen at least 50 out of 100 times were evaluated for comparison. Univariable prediction models based on either ASA class, the Charlson, Elixhauser, or the Rx Risk V comorbidity indices were fitted for comparison. Odds ratios for the final models were estimated with 95 % confidence intervals.

@@

We included `r nmbr(N_shar)` patients (`r mean(df$P_Sex == "Male") * 100` % males) with a mean age of `r age` years in the derivation cohort from the SHAR (Fig. \@ref(fig:flowchart), left panel). 
`r crude_rate_90d * 100` % (n = `r sum(df$outcome_infection_90d)`) of the patients developed a PJI within 90 days, and `r crude_rate_2y * 100` % (n = `r sum(df$outcome_infection_2y)`) within 2 years (Tab. \@ref(tab:tab1)). The proportion of patients with ASA class III was
`r comorb(df$P_ASA == "III")`; `r comorb(df$CCI_index_quan_original > 0)` had at least one comorbidity according to the Charlson comorbidity index, `r comorb(df$ECI_index_sum_all > 0)` according to Elixhauser, and `r comorb(df$Rx_index_pratt > 0)` according to Rx Risk V. In addition, **XXX** patients from **XXX** were included as an external validation cohort (Fig. \@ref(fig:flowchart), right panel).

@@

Patients with `r excl_factors` were uncommon, wherefore those comorbidities were excluded as potential predictors prior to further modeling. ROC-curves (Fig. \@ref(fig:rocs)) and AUC-values (Fig. \@ref(fig:aucci)) for the main and reduced models were no different. For simplicity, we therefore focus on the reduced models. Estimated probability density curves for patients with and without PJI were partially overlapping, although patients with PJI had on average higher predicted probabilities for such event (Fig. \@ref(fig:sep)).

The reduced model for the prediction of PJI within 90 days in the Swedish cohort included `r coefs_text["90 days"]` (Tab. \@ref(tab:coefs)). Predictors associated with a statistically significant risk increase were **secondary osteoarthritis as the underlying reason for THA surgery, male gender, obesity, and the presence of psoriasis, osteoporosis, Paget's or liver disease**. This reduced model for the predicition of PJI within 90 days had an `r get_auc("90 days", "Reduced model", "2")`, whereas ASA class `r get_auc("90 days", "ASA class")`, the Charlson comorbidity index `r get_auc("90 days", "Charlson")`, the Elixhauser comorbidity index `r get_auc("90 days", "Elixhauser")`, and Rx Risk V `r get_auc("90 days", "Rx Risk V")` had less discriminative ability (Fig. \@ref(fig:rocs) and \@ref(fig:aucci)). The reduced model had good accuracy as seen by the calibration plot (Fig. \@ref(fig:calibration), left panel).

The reduced model for PJI within two years included the same predictors, with the addition of `r coefs_text["2 years"]` (Tab. \@ref(tab:coefs)). Predictors conferring a statistically significant risk increase were those **mentioned as predictive for PJI within 90 days, excluding osteoporosis or Paget's disease, and additionally included an underlying diagnosis of AVN or inflammatory joint disease, and the presence of  CNS disease**. The reduced model for the prediction of PJI within two years had an `r get_auc("2 years", "Reduced model", "2")`, whereas ASA class `r get_auc("2 years", "ASA class")`, the Charlson comorbidity index `r get_auc("2 years", "Charlson")`, the Elixhauser comorbidity index `r get_auc("90 days", "Elixhauser")`, and Rx Risk V `r get_auc("90 days", "Rx Risk V")` again had less discriminative ability (Fig. \@ref(fig:rocs) and \@ref(fig:aucci)). Calibration was good for the reduced model (Fig. \@ref(fig:calibration), right panel).

@@

We found that a multivariable main effects logistic regression model based on some easily identifiable clinical conditions, gender, and BMI was considerably better at predicting PJI within 90 days after THA than models based on either ASA class or the Charlson, Elixhauser or Rx Risk V comorbidity indices. A similarly developed model to predict PJI within two years included similar predictors as the model developed to predict PJI within 90 days, but added `r coefs_text["2 years"]` to the list of predictor variables. This model also performed better than the established comorbidity measures in terms of AUC and calibration.

@@

Some previously mentioned risk factors for developing PJI were not selected during our model development. High ASA class is frequently associated with an increased risk of PJI,[@Berbari2012; @Pulido2008; @Namba2012] as are comorbidity in a broader sense, and, more specifically, the presence of cardiovascular comorbidity.[@Ong2009; @Higuera2011] In contrast, we found that ASA class alone was an insufficient predictor of PJI, and none of the cardiovascular disease categories remained as a predictor variable in our final main model.

Cancer seems to confer an increased risk of PJI in several observational studies,[@Bozic2012; @Everhart2013; @Poultsides2013] and it is one of the defined comorbidities in the ACS NSQIP Surgical Risk Calculator,[@Wingert2016a]
20 but this comorbidity did not reach statistical significance in our models.

@@

Although mentioned as risk factors for surgical site infections after joint arthroplasty,[@Rasouli2014a; @Inacio2015a] prediction models based on the comorbidity indices developed by Charlson or Elixhauser and the RxRisk V index alone, resulted in poor predictive power. This agrees with previous findings describing that these measures of comorbidity rather poorly predict mortality and patient reported outcomes after THA.[@Bulow2017; @Bulow2019]

@@

```{r aucci, fig.cap = "Area Under the Receiver Operation Characteristics Curve (AUC) as a measure of predictive discriminative ability with 95 % bootstrap confidence intervals. The reduced model performs no different than the main model, but both of these perform better than traditional comorbidity measures alone."}

@@

tab_categorization %>%
  ungroup() %>% 
  mutate(
    `Comorbidities by groups` = 
      gsub("Heart infarct", "Myocardial infarction", `Comorbidities by groups`)
  ) %>% 
  knitr::kable(
    caption = "Categorization of comorbidities from the Charlson, Elixhauser and Rx RIsk V comorbidities. (CNS = central nervous system.)")
